Discussion Draft

### 2020-04-15 doc/Discussion.rmd
few more thoughts

#

@@

(also outlined in Hanel et al 2018 -- which connects to typical landscape)

A consequence of sexual dimorphism chromatin organization is that interference strength quantified in base pairs will be sexually dimorphic. There are two aspects that define this relationship. First the density of tethering of loops to the axis is highly conserved (Zick Kleck), meaning variability of the axis length of a given chromosome comes from variation in the loop sizes (Z n K).  Second, the signal of crossover interference is mediated through the chromosome axis or SC (kleck Dernberg). Thus the chromatin organization with larger loops (and shorter axis) will have stronger crossover interference when quantified in base pairs, since more base pairs are sequestered in the chromatin loops per unit of axis or SC. Empirical data confirms that interference strength measured in micrometer along the axis is conserved between sexes while interference strength quantified from genetic map is sexually dimorphic, comes from mice [@petkov2007], humans (Gruhn, Lynn). However, due to lack of relevant cytology data in both sexes it is not clear if the many examples of sexual dimorphism in linkage maps is due to chromatin organization.

>>>>

# Hanel et al predicts -- lower genetic diversity near centers -- away from chromosome ends (this is supported )
#finds that the strenght of telomere is connected to chrm size (positive stronger) -- independant of cenromere

#as first outlined in petkov et al and restated in Haenel et al, -- chromatin organization

@@

As outlined in Hanel et al 2018, sexual dimorphism in chromatin organization translates to interference strength in base pairs. Since more DNA is sequestered into the radial loops per unit of the central axis, in the chromosome organization with longer DNA loops, as seen in males for mice (Lynn, gruhn) and humans (petkov), interference strength measured in base pairs will be stronger compared to female linkage maps. The signal for interference is mediated though the chromosome axis (Zickler Kleckner), interference strength measured in micrometers of is conserved between sexes [@petkov2007].

(hanel makes connections to the typical (telomere-bias) --(stronger interference pushes the crossovers to the ends)

<just outlined a consequence of sexual dimorphism in chromatin organization... but this isn't conserved as a

-- not confirmed - if this is a conserved pattern / 
pattern - (general pattern for linkage maps, but there's a lack of cytological data for both sexes)

-- the role of chromatin organization in recombination variation landscapes -- should be studied/ (needs more support)
